---
title: "CDA Project Background Research, Methods, and Data Summaries"
author: "Joshua Ingram"
date: "11/30/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(RColorBrewer)
library(vcd)
library(VGAM)
lc.loans <- read.csv("C:\\Users\\joshi\\Documents\\CDA Final Project Loan Data\\loan.csv")
```

# Numerical and Graphical Data Summaries

First, let's take a look at what exactly we are working with:

```{r}
head(lc.loans)
ls(lc.loans)
nrow(lc.loans)
ncol(lc.loans)
```

We have:

145 variables and 2,260,668 observations (loans) from 2007-2018

Let's see the possible categories for the loan grade:
```{r}
levels(lc.loans$grade)
```

Here's a table that has the counts for all the loan grades:
```{r}
grade.counts.table <- table(lc.loans$grade)
grade.counts.table
grade.prop.long <- as.data.frame(as.table(prop.table(grade.counts.table)))
colnames(grade.prop.long) <- c("Grade", "Proportion")
sum(grade.counts.table)
```

Just to make sure there are no missing values for loan grades, I summed the counts up and it equals the row count... good!

Proportions of grades:
```{r}
prop.table(grade.counts.table)
```

The average loan amount requested($):
```{r}
boxplot(lc.loans$loan_amnt)
hist(lc.loans$loan_amnt)
mean(lc.loans$loan_amnt)
```

Let's look at the distribution of loan grades:
```{r}
ggplot(data = grade.prop.long, aes(x = Grade, y = Proportion, fill = Grade)) + geom_bar(stat = "identity", width = .8) + geom_bar(stat = "identity", width = .8, color = "black", show.legend = F, alpha = 1) + labs(y = "Proportion", title = "Distribution of Loan Grades") + scale_fill_manual(name = "Grade", values = brewer.pal(7, "Reds")[1:7])
```



First let's see if there is any difference in the proprotion of loan grades give an application type:

The number of single borrower loans and joint loans
```{r}
nrow(lc.loans[which(lc.loans$application_type == "Individual"),])
nrow(lc.loans[which(lc.loans$application_type == "Joint App"),])
```

```{r}
grades.apptype.table <- table(lc.loans$application_type, lc.loans$grade)
grades.apptype.matrix <- as.matrix(grades.apptype.table)
grades.apptype.prop <- prop.table(grades.apptype.table, 1)
grades.apptype.long <- as.data.frame(as.table(grades.apptype.prop))
colnames(grades.apptype.long) <- c("Type", "Grade", "Proportion")

ggplot(data = grades.apptype.long, aes(x = Type, y = Proportion, fill = Grade)) + geom_bar(stat = "identity", width = .8) + geom_bar(stat = "identity", width = .8, color = "black", show.legend = F, alpha = 1) + labs(y = "Proportion", title = "Conditional Distribution of Loan Grades by Application Type") + scale_fill_manual(name = "Grade", values = brewer.pal(7, "Reds")[1:7])
```

```{r}
apptype.test <- chisq.test(grades.apptype.matrix)

apptype.test$expected
grades.apptype.matrix
st.resid <- round(apptype.test$stdres, 3)
st.resid
```


Now that we looked at the application types and its effects on loan grades, let's take a look at homeownership status and loan grades
```{r}
lc.loans$home_ownership <- factor(lc.loans$home_ownership, levels = c("OWN", "MORTGAGE", "RENT", "OTHER", "ANY", "NONE"))
grades.ownership.table <- table(lc.loans$home_ownership, lc.loans$grade)
grades.ownership.table

grades.ownership.matrix <- as.matrix(grades.ownership.table)

grades.ownership.prop <- prop.table(grades.ownership.table, 1)
grades.ownership.prop

grades.ownership.long <- as.data.frame(as.table(grades.ownership.prop))
colnames(grades.ownership.long) <- c("Ownership", "Grade", "Proportion")

ggplot(data = grades.ownership.long, aes(x = Ownership, y = Proportion, fill = Grade)) + geom_bar(stat = "identity", width = .8) + geom_bar(stat = "identity", width = .8, color = "black", show.legend = F, alpha = 1) + labs(y = "Proportion", title = "Conditional Distribution of Loan Grades by Home Ownership") + scale_fill_manual(name = "Grade", values = brewer.pal(7, "Reds")[1:7])
```


```{r}
ownership.test <- chisq.test(grades.ownership.matrix)
ownership.test

round(ownership.test$stdres,3)
```

