---
title: "LM HW 4"
author: "Joshua Ingram"
date: "3/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}

```

# Problem 1

## Generating x values
```{r}
set.seed(1)
x <- runif(100, min = -50, max = 50)
```

## Model fitting and visualizations
```{r}
eps <- rnorm(100, 0, 40)
y <- 3 + (2 * x) + eps

lm_fit <- lm(y ~ x)
summary(lm_fit)

plot(y ~ x)
abline(lm_fit)
```

## 1000 simulations
```{r}
set.seed(1)
beta_hats <- numeric(1000)
beta_ci <- matrix(0, nrow = 1000, ncol = 2)
alpha_hats <- numeric(1000)
alpha_ci <- matrix(0, nrow = 1000, ncol = 2)

plot(y ~ x)

for (i in 1:1000){
  
  eps <- rnorm(100, 0, 40)
  
  y <- 3 + (2 * x) + eps
  
  ## Fit least squares regression y ~ x, plot the resulting fit.
  lm_fit <- lm(y ~ x)
  
  conf.ints <- confint(lm_fit)
  #plot(y ~ x)
  abline(lm_fit)
  
  beta_hats[i] <- lm_fit$coefficients[2]
  
  beta_ci[i,] <- conf.ints[2,]
  
  alpha_hats[i] <- lm_fit$coefficients[1]
  
  alpha_ci[i,] <- conf.ints[1,]
}

# overlaying the population line
abline(3, 2, lw = 3, col = "red")
```

## Lab with beta estimations

### Practical and Theoretical Comparisons

#### Expected Values

Practical:
```{r}
mean(beta_hats)
```

Theoretical:

$E[\hat{\beta}] = \beta = 2$

These two are very close (the practical value is .002 off).

#### Variances

Practical:
```{r}
var(beta_hats)
```

Theoretical:

$V[\hat{\beta}] = \frac{\sigma^2}{\Sigma_i(x_i - \bar{x})^2}$
```{r}
theor_var_beta <- 1600/sum((x - mean(x))^2)
theor_var_beta
```

The practical and theoretical values are virtually the same.

#### Distribution

Practical:
```{r}
hist(beta_hats, freq = F)
```


Theoretical:
$\hat{\beta} \text{ ~ } N(\beta, V[\hat{\beta}])$
```{r}
hist(beta_hats, freq = F)
my.dnorm <- function(z) dnorm(z, 2, sqrt(theor_var_beta))
curve(my.dnorm,
      from = 1.4, to = 2.6,
      add = T, col = "red")
```
Theoretical distribution is in red... histograms are the practical distribution... which very closely follows a normal distribution.

#### Confidence Interval Coverage

Practical:
```{r}
mean(beta_ci[,1] < 2 & beta_ci[,2] > 2)
```

Theoretical:
0.95

The practical estimate is .007 off, but extremely close to 95% coverage.


## Lab with alpha estimations

### Practical and Theoretical Comparisons

#### Expected Values

Practical:
```{r}
mean(alpha_hats)
```


Theoretical:
$E[\hat{\alpha}] = \alpha = 3$

Very close in values

#### Variances

Practical:
```{r}
var(alpha_hats)
```


Theoretical:
$V[\hat{\beta}] = \frac{\sigma^2 \Sigma_ix_i^2}{n\Sigma_i(x_i - \bar{x})^2}$
```{r}
theor_var_alpha <- (1600 * sum(x^2))/(length(x)*sum((x - mean(x))^2))
theor_var_alpha
```

Theoretical variance and practical variance are very close in value as expected.

#### Distribution

Practical:
```{r}
hist(alpha_hats)
```


Theoretical:
$\hat{\alpha} \text{ ~ } N(\alpha, V[\hat{\alpha}])$
```{r}
hist(alpha_hats, freq = F)
my.dnorm <- function(z) dnorm(z, 2, sqrt(theor_var_alpha))
curve(my.dnorm,
      from = -10, to = 15,
      add = T, col = "red")
```

Both follow the expected distributions.

#### Confidence Interval

Practical:
```{r}
mean(alpha_ci[,1] < 2 & alpha_ci[,2] > 2)
```


Theoretical:
0.95

Practical is only .002 off, so very close.
