---
title: "Homework_3"
author: "Joshua Ingram"
date: "9/17/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nnet)
library(car)
library(effects)
library(AER)
library(pscl)

data("NMES1988")
ornstein <- read.csv("c:/main/projects/undergraduate-coursework/Topics in Statistical Inference for Data Science/Datasets/Ornstein.txt", sep="")
nmes <- NMES1988
```

# Problem 1

## 1.

***Note:*** I am using $\lambda$ instead of $\mu$ for my notation in this homework. I've been doing other work with the poisson distribution and am using $\lambda$ for notation for that. It's just easier to stay consistent all around. Let me know if I need to stick with $\mu$ after this.

$$
Y_i \sim_{ind.} Pois(\lambda), \; i = 1, 2,...,248, \text{ where } \lambda \text{is the average number of interlocks}
$$
$$
log(\lambda_i) = \beta_0 + \beta_1(assets_i) + \beta_2I_{nationOTH, i} + \beta_3I_{nationUK,i} + \beta_4I_{nationUS,i} +
$$
$$
\beta_5(I_{nationOTH, i} * assets_i) + \beta_6(I_{nationUK, i} * assets_i) + \beta_7(I_{nationUS, i} * assets_i)
$$
$$
\text{CAN is baseline category, } I_{nationOTH} \in \{0= not \; other \; foreign, 1=other \; foreign\}
$$
$$
I_{nationUK} \in \{0=not \; UK, 1=UK\}, \; I_{nationUS} \in \{0=not \; US, 1=US\}
$$


## 2.

```{r}
fit_1 <- glm(interlocks ~ assets + nation + nation:assets, family = poisson, data = ornstein)
summary(fit_1)
```

Test for Interaction:

$H_0: \; \beta_5 = \beta_6 = \beta_7 = 0$

$H_A: \; \{\exists \; \beta_{j} \neq 0 \; | \;  j = 5,6,7\}$

$\alpha = 0.05$

Null Model: $Y_i \sim_{ind.} Pois(\lambda), \; i = 1, 2,...,248, \; log(\lambda_i) = \beta_0 + \beta_1(assets_i) + \beta_2I_{nationOTH, i} + \beta_3I_{nationUK,i} + \beta_4I_{nationUS,i}$

LRT statistic = $2log(\frac{L_1}{L_0} ) = G^2_0 \sim \chi^2_3$

p-value: $P(\chi^2_{3} \geq G^2_0)$

```{r}
fit_null <-  glm(interlocks ~ assets + nation, family = poisson, data = ornstein)
anova(fit_null, fit_1, test = "LRT")
```

After performing the likelihood ratio test, we get an extremely small p-value (basically 0). We have significant evidence to reject the null hypothesis and our interaction term is statisically significant.

## 3.

```{r}
1- summary(fit_1)$deviance/summary(fit_1)$null.deviance
```

$R^2$ = 0.433715

43.4% of the variation in our response, the number of interlocks, is explained by our model.

## 4.

For a firm with the U.S. as the nation of control, per 1 million dollar increase in assets, the average number of interlocks will increase by a factor of $e^{7.647e-7}$. If the U.S. controlled firm has 0 dollars in assets, the average number of interlocks will be $e^{1.072e+00}$ time LESS than a Canadian controlled firm.

*I'm not specifying "ceteris paribus" since nation and assets are our only two variables in the model and these are being explicitly addressed in the interpretation.*

# Problem 2

## 1.

```{r}
fit_2 <- glm(visits ~ chronic + age + gender + income + insurance, family = poisson, data = nmes)
summary(fit_2)
```

### a.

Yes, there is evidence of overdispersion because the residual deviance is much greater than the residual degrees of freedom. This means that our variance is greater than expected, which under a poisson model, should be the same as the mean.

### b.

We should use a Quasi-Poisson model.

```{r}
fit_quasi <- glm(visits ~ chronic + age + gender + income + insurance, family = quasipoisson, data = nmes)
summary(fit_quasi)
```

Differences:

The standard errors are greater for the quasi-poisson, thus affecting the t-statistics and p-values given for each beta. The inverse is true for the regular poisson model.

Similarities:

Both the quasi-poisson model and the poisson model have the same exact estimates for the betas and the Null/Residual deviances are the same (and degrees of freedom). 

### c.

```{r}
Anova(fit_2)
Anova(fit_quasi)
```

With the quasi-poisson model I would drop the age and incoe predictors, but for the regular poisson model I would drop only the income predictor.

## 2.

### a.

By looking at the distribution of counts of interlocks, we may be running into an issue of having too many zero-counts of interlocks. It may be more appropriate to fit a zero-inflated poisson model to our data.

### b.

Full GLM Model Formulation:

$$
p_i = P(Y_i \in \; "no \; visists)
$$
$$
log(\frac{p_i}{1-p_i}) = \gamma_0 + \gamma_1chronic_i + \gamma_2age_i + \gamma_3I_{gender,i} + \gamma_4income_i + \gamma_5I_{insurance,i}
$$
$$
Y_i \sim_{ind.} Pois(\lambda_i), \; \text{where } \lambda \text{ average number of physician visits}
$$
$$
log(\lambda_i) \beta_0 + \beta_1chronic_i + \beta_2age_i + \beta_3I_{gender,i} + \beta_4income_i + \beta_5I_{insurance,i}
$$
$$
I_{gender} \in \{0=female, 1=male\}, \; I_{insurance} \in \{0=no, 1=yes\}
$$

### c.

```{r}
fit_3 <- zeroinfl(visits ~ chronic + age + gender + income + insurance, data = nmes)
summary(fit_3)
Anova(fit_3)
```

We use the Anova() function to test the predictor as a whole. Although the name was not explicitly given in class, the test statistic follows a chi-square distribution.

For age:

$H_0: \gamma_2 = \beta_2 = 0$

$H_0: \{\exists \; \beta_{2} \; or \; \gamma_2 \neq 0 \}$

$\alpha = 0.05$

income is statistically insignificant based on the output from Anova().

### d.

```{r}
fit_4 <- zeroinfl(visits ~ chronic + age + gender + insurance, data = nmes)
summary(fit_4)
```

Chronic:

logit - For every one additional chronic conditions, the odds of an individual not having any physician visits decrease by a factor of $e^{0.554}$, ceteris paribus.

poisson - For every one additional chronic condition, the average number of physician visits will increase by a factor of $e^{0.154}$, ceteris paribus.

Insurance:

logit - For a person with insurance, the odds of them having no physician visits is $e^{0.92}$ times lower than those that have no insurance, ceteris paribus.

poisson - For a person with insurance, the average number of physician visits is $e^{0.11}$ times greater than those without insurance, ceteris paribus.