---
title: "HW1"
author: "Joshua Ingram"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ISwR)
```

# DUE: Wednesday, September 11th, at 5pm via Canvas Submission.


 

# Problem #1

Indicate whether each scenario is a classification or regression problem, and how many variables are used to describe an observation.
    
  1. Predicting if a student is going to SAT the college-level introductory statistics course based on their overall high school GPA, their algebra grade and their geometry grade.
  
  Answer: Classification problem with 3 variables used to describe the observation.
  
  2. Same as before, only now predicting the letter grade (A,B,C,D,F).
  
  Answer: Classification problem with 3 variables used to describe the observation.
  
  3. Predicting score differential in a football game between **two opponents**. We assume that it is mid-season, and obtain the data on the games completed by then, studying such variables as: yards (gained & allowed), touchdowns (scored & allowed), turnovers (committed & forced), whether it rains or not.
  
  Answer: Regression problem with 4 variables used to describe the observation.
  
  4. Forecasting the total number of items sold by a Walmart supermarket based on the day of the week, location and store size.

  Answer: Regression problem with 3 variables used to describe the observation.

# Problem #2

In data set $cystfibr$ of $ISwR$ package, proceed to:

```{r}
head(cystfibr)
```


1. Use a looping construct to loop through each variable and calculate mean and standard deviation *without* using $R$'s customized functions like $mean(), sd()$.



Mean and SD
```{r}

for (var in 1:ncol(cystfibr)){
  
  name <- colnames(cystfibr[var])
  len <- length(cystfibr[[name]])
  mean <- sum(cystfibr[[name]]) / len
  
  print(name)
  print(paste("Mean =", mean))
    
  sum1 <- 0
    
  for (i in cystfibr[[name]]){
      
    sum1 <- sum1 + (i - mean)^2
      
  }
    
  sd <- sqrt(sum1/(len-1))
  print(paste("Standard Deviation =", sd))

}
  
```


2. Create a matrix with rows corresponding to variables, columns - to calculated summaries (mean, sd), and fill it out with the results of the previous step accordingly. Please assign appropriate row  and column names as well.

```{r}

matrix1 <- matrix(c("name", "mean", "sd"), ncol = 3)

for (var in 1:ncol(cystfibr)){
  
  vec <- c()
  
  name <- colnames(cystfibr[var])
  len <- length(cystfibr[[name]])
  mean <- sum(cystfibr[[name]]) / len
  
  vec[1] <- name
  vec[2] <- mean
    
  sum1 <- 0
  
  for (i in cystfibr[[name]]){
      
    sum1 <- sum1 + (i - mean)^2
      
  }
    
  sd <- sqrt(sum1/(len-1))
  
  vec[3] <- sd
  
  matrix1 <- rbind(matrix1, vec)

}

matrix1
```


3. Proceed to:
    a. Extract all information on first $10$ observations.
```{r}
head(cystfibr, 10)
```
    
    b. Extract all information on observations $\# 5,10 \ \& \ 15$.
```{r}
cystfibr[c(5, 10, 15),]
```

    c. Extract all information on last $10$ observations.
```{r}
tail(cystfibr, 10)
```

    d. Extract all information on the first *variable*.
```{r}
cystfibr[,1]
```

    e. Extract information on the first *variable* for observations $\# 5,10,15$.
```{r}
cystfibr[c(5, 10, 15), 1]
```

    f. Extract all information on the patients of above-average height.
```{r}
hmean <- mean(cystfibr$height)
subset(cystfibr, height > hmean, select = height)
```
    
4. Show all the information on 5 oldest patients.
```{r}
cystfibr[tail(order(cystfibr$age), 5),]
```


5. Using a $subset()$ function, only select the observations that have above-average values for both $weight$ and $height$.
```{r}
subset(cystfibr, weight > mean(cystfibr$weight) & height > mean(cystfibr$height))
```


6. Make a scatterplot for an arbitrary pair of quantitative variables, comment on the relationship.
```{r}
plot(tlc ~ weight, data = cystfibr)
```

Plotting weight and total lung capacity, with TLC as the response, there may be a slightly noticeable relationship, with tlc decreasing as weight increases. However, it may not be a very strong relationship.

# Problem #3

Define your own function in $R$ and demonstrate its functionality:

   1. Function needs to contain:
      a. at least one assignment operation ('<-')
      b. at least one $required$ argument (without a default value)
      c. at least one argument with a default value
      d. at least one conditional $if$-expression
   2. Demonstrate:
      a. 3 calls to show that function works properly on various $required$ inputs}
      b. a call in which you specify a non-default value

Function idea examples:  check if a string/sequence has a certain property, or make a cyclic algebraic calculation,  or provide various plots for inputted sequence/data... 

**Example**: Function $is.increasing()$ allows to check if a sequence of numbers is monotonically increasing (returns TRUE) or not (returns FALSE). Upon user's request, it can also print out the consecutive differences.

```{r}
is.increasing <- function(x,printout=FALSE){
  n <- length(x)
  dif.vec <- x[2:n]-x[1:(n-1)]
  if (printout == TRUE) print(dif.vec)
  if (sum(dif.vec<=0) != 0) return(F)  # Important note: function stops after executing the
  return(T)                            # first 'return' statement it gets to
}

x <- c(1:10)
is.increasing(x)

y <- c(5,5,5)
is.increasing(y)

z <- c(1,5,3,7,9)
is.increasing(z)

is.increasing(z,printout=T)   # The call where you specify the non-default value
```


My Function (gives the user the mean and/or standard deviation of a vector of numbers... func = 1 gives mean, func = 2 gives sd, func = 3 gives both sd and mean):
```{r}
sd.mean <- function(vector, func = 3){
    length <- length(vector)
    mean <- sum(vector)/length
      
    sum <- 0
      
    for (i in vector){
      sum <- sum + (i - mean)^2
    }
    
    sd <- sqrt(sum/(length-1))
      
  if (func == 1) print(paste("Mean =", mean))
    
  if (func == 2) print(paste("Standard Deviation =", sd))
    
  if (func == 3) print(paste("Mean =", mean, "Standard Deviation =", sd))
  
}

sd.mean(c(1,4,5,6,12,9))

sd.mean(c(93,123,45,199,33,55))

sd.mean(c(55,66,223,44,6,22))

sd.mean(c(1,2,4,6,2,7,8), func = 1)
```

# Problem #4

Count the total number of boys in data set $juul$ (from package $ISwR$), who have insulin-like growth factor greater than 400, in three different ways (make sure the output is the same though):
```{r}
head(juul, 15)
```

* Using for-loop
```{r}
count <- 0

for (i in 1:nrow(juul)){
  if (juul$sex[i] == 1 & is.na(juul$sex[i]) != TRUE) {
    if (juul$igf1[i] > 400 & is.na(juul$igf1[i]) != TRUE){
      count = count + 1
    }
  }
}
  
print(count)

```

* Using vectorized operations (can be done in one line)
```{r}
length(which(juul$igf1 > 400 & juul$sex == 1))
```

* Using data frame subsetting (can be done in one line)
```{r} 
nrow(juul[juul$sex == 1 & is.na(juul$sex) != TRUE & is.na(juul$igf1) != TRUE & juul$igf1 > 400,])
```


