---
title: "SL Homework 3"
author: "Joshua Ingram"
date: "September 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
advertising <- read.csv("C:/Users/joshi/Downloads/Advertising.csv")
```

## Problem 1

```{r}
head(advertising, 6)
sales <- advertising$sales
tv <- advertising$TV
radio <- advertising$radio
newspaper <- advertising$newspaper

plot(sales ~ tv)

```

#### 1.

$\beta_0$ and $\beta_1$ estimates:
```{r}
tv_mean <- mean(tv)
sales_mean <- mean(sales)
beta_hat_1 <- 0
beta_1_numer <- 0
beta_1_denom <- 0

for (i in 1:length(sales)){
  beta_1_numer <- beta_1_numer + (tv[i] - tv_mean) * (sales[i] - sales_mean)
  beta_1_denom <- beta_1_denom + (tv[i] - tv_mean)^2
}

beta_hat_1 <- (beta_1_numer / beta_1_denom)
print("beta_1 estimate")
beta_hat_1

beta_hat_0 <- sales_mean - (beta_hat_1*tv_mean)
print("beta_0 estimate")
beta_hat_0

print("Values from lm()")
lm(sales~tv)
```

Standard Errors of estimates:
```{r}
s_num <- 0
n <- length(sales)

for (i in 1:length(sales)){
  s_num <- s_num + (sales[i] - (beta_hat_0 + (beta_hat_1*tv[i])))^2
}

s <- sqrt(s_num/(n-2))

denom <- 0

for (i in 1:length(sales)){
  denom <- denom + (tv[i] - tv_mean)^2
}

se_beta_hat_0 <- s * sqrt((1/n)+(tv_mean^2/denom))
print("se beta_0")
se_beta_hat_0

se_beta_hat_1 <- s/sqrt(denom)
print("se beta_1")
se_beta_hat_1

print("Values from lm()")
summary(lm(sales~tv))
```


#### 2.
```{r}
plot(sales ~ tv, ylab = "Sales", xlab = "TV", col = "red", pch = 16)
abline(lm(sales~tv), col = "blue")
segments(tv, sales, tv, predict(lm(sales~tv)))
```


## Problem 2

```{r}
set.seed(2)

n.rep <- 1000
conf_int_b0 <- matrix(0, nrow=n.rep, ncol=2)
conf_int_b1 <- matrix(0, nrow=n.rep, ncol=2)

for (r in 1:n.rep){
  X <- rnorm(200, mean=0, sd=1)
  eps <- rnorm(200, mean=0, sd=4)
  Y <- 2 + 3*X + eps
  lm <- lm(Y~X)
  conf <- confint(lm)
  conf_int_b0[r,] <- conf[1,]
  conf_int_b1[r,] <- conf[2,]
}


print(mean(conf_int_b0[,1] < 2 & 2 < conf_int_b0[,2]))
print(mean(conf_int_b1[,1] < 3 & 3 < conf_int_b1[,2]))

```

These confidence intervals contain our betas roughly 95% of the time, as expected.


## Problem 3

#### Sales onto Radio
```{r}
lm.radio <- lm(sales~radio)
summary(lm.radio)

```

1.
HO: $\beta_{1}$ = 0
HA: $\beta_{1}$ =/= 0
$\alpha$ = .05
```{r}
summary(lm.radio)
```
With alpha at .05, we received a p-value of 2.2e-16 which gives us significant evidence to reject the null hypothesis in favor of the alternative that states that $\beta_{1}$ is not equal to 0. This gives us enough evidence to believe that there is some relationship between the predictor and the response.

2.
Intercept: 9.31
Slope: .2025
Our intercept of 9.31 tells us when there is $0 invested in radio, sales should be at about 9,310 items sold.
Our slope tells us that for every $1,000 invested in readio, there should be 202.5 more items sold.

3.
```{r}
confint(lm.radio)
```
We are 95% confident that when spending $1,000 more on radio there will be between 162.24 and 242.74 additional units sold

4.
```{r}
prediction.radio <- 9.31 + (.2025 * 20)
prediction.radio
pred.frame <- data.frame(radio=seq(min(radio), max(radio), length=100))
pp <- predict(lm.radio, int="p", newdata=pred.frame)
pc <- predict(lm.radio, int="c", newdata=pred.frame)
plot(radio ,sales, main="Fitted Regression Line with Bands.",     
     ylim=range(sales, pp, na.rm=T))
matlines(pred.frame$radio, pc, lty=c(1,2,2), col=2)
matlines(pred.frame$radio, pp, lty=c(1,3,3), col=4)
legend("topleft",legend = c("Confidence","Prediction"), col=c(2,4), lty=c(2,3))
```
Based on our model, if we invested $20,000 into radio we would predict for there to be 13,360 items sold.
The prediction bands are wider than the confidence bands because the confidence bands are the interval for which the average prediction can be found 95% of the time, whereas the prediction bands are wider because they contain 95% of the values of the observations.

5.
Residual Standard Error (from summary()): 4.275
The standard deviation of the residuals is 4.275

6.
R^2 = .332
33.2% of variance in sales is explained by radio advertisement spending

#### Sales onto Newspaper
```{r}
lm.newspaper <- lm(sales~newspaper)
summary(lm.newspaper)
```

1.
HO: $\beta_{1}$ = 0
HA: $\beta_{1}$ =/= 0
$\alpha$ = .05
```{r}
summary(lm.newspaper)
```
With alpha at .05, we received a p-value of .001148 which gives us significant evidence to reject the null hypothesis in favor of the alternative that states that $\beta_{1}$ is not equal to 0. This gives us enough evidence to believe that there is some relationship between the predictor and the response.

2.
Intercept: 12.35141
Slope: .05469
Our intercept of 12.35141 tells us when there is $0 invested in radio, sales should be at about 12,351 items sold.
Our slope tells us that for every $1,000 invested in readio, there should be 54.69 more items sold.

3.
```{r}
confint(lm.newspaper)
```
We are 95% confident that when spending $1,000 more on newspaper there will be between 22 and 83.4 additional units sold

4.
```{r}
prediction.newspaper <- 12.3514 + (.05469 * 20)
prediction.newspaper
pred.frame <- data.frame(newspaper=seq(min(newspaper), max(newspaper), length=100))
pp <- predict(lm.newspaper, int="p", newdata=pred.frame)
pc <- predict(lm.newspaper, int="c", newdata=pred.frame)
plot(newspaper,sales, main="Fitted Regression Line with Bands.",     
     ylim=range(sales, pp, na.rm=T))
matlines(pred.frame$newspaper, pc, lty=c(1,2,2), col=2)
matlines(pred.frame$newspaper, pp, lty=c(1,3,3), col=4)
legend("topleft",legend = c("Confidence","Prediction"), col=c(2,4), lty=c(2,3))
```
Based on our model, if $20,000 was invested into newspaper we would predict for there to be 13,445 items sold.
The prediction bands are wider than the confidence bands because the confidence bands are the interval for which the average prediction can be found 95% of the time, whereas the prediction bands are wider because they contain 95% of the values of the observations.

5.
Residual standard error (from summary()): 5.092
The standard deviation of the residuals is 5.092

6.
R^2 = .05212
5.2% of variance in sales is explained by newspaper advertisement spending
