---
title: "CDA HW 6"
author: "Joshua Ingram"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
yawn <- read_csv("C:/Users/joshi/Downloads/Yawn.csv")
```

# Yawn Study

## a.

$logit(\pi_i) = log(\frac{\pi_i}{1 - \pi_i}) = \alpha + \beta x_i$

$\frac{\pi_i}{1 - \pi_i} = e^{\alpha + \beta x_i}$

$(1-\pi_i)e^{\alpha + \beta x_i} = \pi_i$

$e^{\alpha + \beta x_i} - \pi_ie^{\alpha + \beta x_i} = \pi_i$

$e^{\alpha + \beta x_i} = \pi_i(1 + e^{\alpha + \beta x_i})$

$\pi_i= \frac{e^{\alpha + \beta x_i}}{1 + e^{\alpha + \beta x_i}}$

## b.

$\pi_i= \frac{e^{\alpha + \beta x_i}}{1 + e^{\alpha + \beta x_i}}$

$\pi_i = \frac{e^{\alpha + \beta x_i}}{e^{\alpha + \beta x_i}(\frac{1}{e^{\alpha + \beta x_i} + 1})}$

$\pi_i = \frac{1}{1 + e^{-\alpha - \beta x_i}}$

## c.

$logit(\pi_i) = \alpha + \beta$

$\pi_i= \frac{e^{\alpha + \beta}}{1 + e^{\alpha + \beta}}$

## d.

$logit(\pi_i) = \alpha$

$\pi_i= \frac{e^{\alpha}}{1 + e^{\alpha}}$

## e.

```{r}
head(yawn)
glm.obj <- glm(Yawn ~ Age, family = binomial, data = yawn)
summary(glm.obj)
```


## f.

MLE of \alpha = 2.20772

MLE of \beta = -0.04418

## g.

```{r}
2.20772 - 0.04418*22
```

log-odds = 1.23576

## h.

```{r}
2.71828^(2.20772 - 0.04418*22)
```

odds = 3.44099

## i.

```{r}
1/(1 + 2.71828^(-2.20772 + 0.04418*22))
```

probability = 0.774825