---
title: "CDA HW 4"
author: "Joshua Ingram"
date: "October 2, 2019"
output: html_document
---

```{r setup, include=FALSE}

```


# 1. Likelihood for a random sample of size n from a Poisson distribution

a.
L(λ;y1,...,yn) = $\prod_{i=1}^n$ ($\lambda^{y_i}$ $e^{-\lambda}$)/$y_i$!

b.
l(λ;y1,...,yn) = $\sum_{i=1}^n$ $y_i$log($\lambda$) $-\lambda$ - log($y_i$!)

c.
```{r}
lambda <- c(seq(0,25,0.01))
yi <- c(3, 0, 4, 3, 2, 4, 1, 5)
likelihood <- c()
logs <- c()
n <- length(yi)

for (i in lambda){
  sum <- 0
  logs <- c()
  for (j in 1:8){
    sum <- (yi[j]*log(i)) -i - log(factorial(yi[j]))
    logs <- append(logs, sum)
  }
  likelihood <- append(likelihood, sum(logs))
}
plot(likelihood~lambda)

```


d.
```{r}
mle<-lambda[match(max(likelihood, na.rm = TRUE), likelihood)]
mle
```

MLE is 2.75

# 2. Statistical Inference about the Poisson mean λ.

a. 
.....

b.
For a large sample size n, the MLE follows a normal distribution with a mean of lambda and SD of sqrt(lambda/n)

c.
```{r}
null <- 2
z.wald <- (mle - null) / sqrt(mle/n)
z.wald
pnorm(mean = null, sd = sqrt(null/n), q = z.wald) * 2
```
test statistic = 1.2792
p-value = .1494

d.
```{r}
z.score <- (mle - null) / sqrt(null/n)
z.score
pnorm(mean = null, sd = sqrt(null/n), q = z.score) * 2
```
test statistic = 1.5
p-value = .317

e.
```{r}
yi <- c(3, 0, 4, 3, 2, 4, 1, 5)
logs1 <- c()
sum <- 0

for (i in yi){
  sum <- (yi[i]*log(2.75)) -2.75 - log(factorial(yi[i]))
  logs1 <- append(logs1, sum)
}

logs0 <- c()
sum <- 0

for (i in yi){
  sum <- (yi[i]*log(2)) -2 - log(factorial(yi[i]))
  logs0 <- append(logs0, sum)
}

l1 <- sum(logs1)
l0 <- sum(logs0)
ratio <- 2 * (l1 - l0)
ratio
pnorm(mean = null, sd = sqrt(null/n), q = ratio) * 2
```
likelihood ratio test statistic = 1.601
p-value = .425

f.
```{r}
conf.int <- mle + c(-1,1) * (1.96 * sqrt(mle/n))
conf.int
```
confidence interval: (1.6, 3.899)
