---
title: 'Intro to Categorical: Homework 1'
author: "Joshua Ingram"
date: "August 29, 2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(qcc)

alligator.data <- read.csv("https://sites.williams.edu/bklingen/files/2018/09/alligatorfood.csv")
```

# 2. Alligator Food Choice

### a. Frequency Table

```{r}
gator.table <- summary(alligator.data$food)
gator.table
```


### b. Bar Chart
```{r}
barplot(gator.table, main = "Alligator Food Choice", ylim = c(0, 100), xlab = "Food Choice", ylab = "Count", col = c("green1", "blue1", "green3", "blue3", "green4"))
```


### c. Pareto Chart
```{r}
pareto.chart(gator.table, ylab = "Counts", main = "Alligator Food Choice")
```

### d. Pie Chart
```{r}
pie(gator.table, main = "Alligator Food Choice")
```

### e. Distribution Explanation

Based on the distribution of food choices from this dataset, alligators seemed to have eaten more fish than any other category (43% of the total counts), followed by invertebrate (28% of the counts). Birds were the least common choice, being only 6% of the total counts of food choices.
