---
title: "CDA Homework 2"
author: "Joshua Ingram"
date: "September 10, 2019"
output: html_document
---

```{r setup, include=FALSE}

```


Questions:

1.1: In the following examples, identify the natural response variable and the explanatory variables.

a.
Response: Attitude toward gun control
Explanatory: gender, mather's education

b.
Response: Heart Disease
Explanatory: Blood Pressure, Cholesterol Level

c.
Response: Vote for President
Explanatory: Race, Religion, Annual Income


1.2 Which scale of measurement is most appropriate for the following variables - nominal or ordinal?

a.
nominal

b.
ordinal

c.
ordinal

d.
nominal

e.
nominal

f.
ordinal



1.3 Each of the 100 multiple-choice questions on an exam has four possible answers but one correct response. For each question, a student randomly selects one response as the answer.

a.
Binomial Distribution

b.
It would be surprising if a student recieved 50 correct responses, as the mean is 25 (.25 * 100 (pi x n)) and the standard deviation is 4.33 (sqrt(n(pi)(1-pi))). For a student to get 50 correct responses after guessing randomly, it would be more than 3 standard deviations away from the mean, which is extremely unlikely based upon the empirical rule.

i.
```{r}
dbinom(50, 100, .25)
```


ii.
```{r}
pbinom(49, 100, .25)
```


iii.
```{r}
1 - pbinom(49, 100, .25)
```



1.4 In a particular city, the population proportion PI supports an increase in the minimum wage. For a random sample of size 2, let Y = number who support an increase.

a.
```{r}
y0 <- dbinom(0, 2, .50)
print(paste('P(Y=0) =', y0))
y1 <- dbinom(1, 2, .50)
print(paste('P(Y=1) =', y1))
y2 <- dbinom(2, 2, .50)
print(paste('P(Y=2) =', y2))

mean.dist <- .50 * 2
sd.dist <- sqrt(2*.5*(1-.5))
print(paste("The mean of the distribution is", mean.dist, "and the standard deviation is", sd.dist))
```


b.
Although we do not know $\pi$, based on the observed value of y = 1 from a sample size of 2, we know that the ML estimate, $\hat($\pi$)$ = y/n, which is 1/2 = .50.

Distribution:
```{r}
pi = seq(0,1,0.01)
like.pi = dbinom(1, 2, pi)
plot(pi, like.pi, type="l")
```

the ML estimate pi-hat is .50 as the ML, as seen on the graph, is at pi =.50 (where ML is maximum)... also y/n = .50

1.6 Genotypes AA< Aa, and aa occur with probabilities (PI_1, PI_2, PI_3). For n = 3 independent observations, the observed frequencies are (y_1, y_2, y_3).

a.
when obtaining the 3 observed frequencies over n=3 observations, we can deduce the value of y_3 from knowing that y_1 + y_2 + y_3 = n, thus y_3 = n - y_1 - y_2. So if we know n, y_1, and y_2, then we know y_3.

b.
possible outcomes:
(AA,AA,AA) (AA,AA,Aa) (AA,AA,aa) (AA,Aa,Aa) (AA,Aa,aa) (AA,aa,aa) (Aa,Aa,Aa) (Aa,Aa,aa) (Aa,aa,aa) (aa,aa,aa)

```{r}
grid = expand.grid(0:3,0:3,0:3)
index = rowSums(grid)
index1 = index == 3
grid3 = grid[index1,]
multinomdist <- apply(grid3,1,dmultinom,size=3,prob=c(.25,.50,.25))
multinomdist1 <- cbind(grid3, multinomdist)
multinomdist1

print("Checking to see if probabilities add up to 1")
sum(multinomdist1[,4])
```



c.
The probability distribution would be a Binomial Distribution for y_1 alone.


1.7

a.
```{r}
dbinom(0, 6, (5/6))
```

